% Chapter 1

\chapter{Conclusion} % Main chapter title

\label{Conclusion} % For referencing the chapter elsewhere, use \ref{Chapter1} 

%----------------------------------------------------------------------------------------

% Define some commands to keep the formatting separated from the content 

%----------------------------------------------------------------------------------------

\section{Summary of Thesis}
We used classification technique to discover knowledge from biis student data in this thesis project. All programming and necessary statistical analysis was done using java programming. The raw data set was full of impurities. So, we have to do an extensive amount of preprocessing on the data before applying it in our algorithm. Training data set was prepared very carefully and with logistic reasoning. The results of the classier was statistically analyzed and was stored in tabular form and graphical form for better visualization of the findings. The accuracy of the classifier is about 76\%. The low accuracy is due to the variations and impurities of the input data.   
%----------------------------------------------------------------------------------------

\section{General Findings}
The Classifier has five class labels. They are excellent,good,moderate,poor,very poor. After completing statistical analysis some major findings are
\begin{itemize}
\item EEE and CSE departments have a higher percentage in top levels. ARCH department has a higher percentage at the lowest level.

\item On average female students have better academic performance than male students.
\item Attached students dominate in the higher classes over the resident students.
\item Class Attendance and class attendance is very important for a student's overall performance.Students with good attendance trends to be in higher classes.
\item Class test marks are also very important for a student. Students with high percentage of class test marks trend to be in the higher classes. On the other hand students with lower percentage of class test marks fall in lower classes.

\end{itemize}


%----------------------------------------------------------------------------------------

\section{Future Works}

We used only classification rules to discover knowledge from data. In future we want to combine the clustering algorithm  with this one to get more accurate result.
\\The current accuracy is low. So we want to improve the accuracy by applying accuracy improvement algorithm.
\\We want to use this project to make a complete system for student evaluation.
  

%----------------------------------------------------------------------------------------

